##### Question 1
What is the variance of the distribution of the average an IID draw of n observations from a population with mean μ and variance σ2.

$$\frac{\delta^2}{n}$$

##### Question 2
Suppose that diastolic blood pressures (DBPs) for men aged 35-44 are normally distributed with a mean of 80 (mm Hg) and a standard deviation of 10. About what is the probability that a random 35-44 year old has a DBP less than 70?

```{r}
z <- abs((70 - 80)/ 10) 
1-pnorm(z)
pnorm(70, mean = 80, sd = 10)
```

##### Question 3
Brain volume for adult women is normally distributed with a mean of about 1,100 cc for women with a standard deviation of 75 cc. About what brain volume represents the 95th percentile?
```{r}
qnorm(0.95, mean = 0, sd = 1) * 75 + 1100
qnorm(0.95, mean = 1100, sd = 75)
```

##### Question 4
Refer to the previous question. Brain volume for adult women is about 1,100 cc for women with a standard deviation of 75 cc. Consider the sample mean of 100 random adult women from this population. Around what is the 95th percentile of the distribution of that sample mean?

```{r}
qnorm(0.95, mean = 1100, sd = 75/sqrt(100))
```

##### Question 5
You flip a fair coin 5 times, about what's the probability of getting 4 or 5 heads?

$$5C4*(1/2)^5 + 5C5*(1/2)^5$$

```{r}
choose(5,4) * (1/2) ^ 5 + choose(5,5) * (1/2) ^ 5
pbinom(3, size = 5, prob = 0.5, lower.tail = FALSE)
```

##### Question 6
The respiratory disturbance index (RDI), a measure of sleep disturbance, for a specific population has a mean of 15 (sleep events per hour) and a standard deviation of 10. They are not normally distributed. Give your best estimate of the probability that a sample mean RDI of 100 people is between 14 and 16 events per hour?

The standard error is $10/\sqrt{100} = 1$

```{r}
z <- 1
1 - pnorm(z)*2
pnorm(16, mean = 15, sd = 1) - pnorm(14, mean = 15, sd = 1)
```

##### Question 7
Consider a standard uniform density. The mean for this density is .5 and the variance is 1 / 12. You sample 1,000 observations from this distribution and take the sample mean, what value would you expect it to be near?

The law of large number 
```{r}
mean(rnorm(1000, mean = 0.5, sd = sqrt(1/12)))
```

##### Question 8
The number of people showing up at a bus stop is assumed to be Poisson with a mean of 5 people per hour. You watch the bus stop for 3 hours. About what's the probability of viewing 10 or fewer people?

```{r}
ppois(10, lambda = 5*3)
```